Update Meeting: 14-07-2021

### 2021-09-15 analysis/update_meeting_05_07_2021.Rmd
update all update meetings Rmds

---
title: "Update Meeting: 05-07-2021"
author: "jennysjaarda"
date: "2021-07-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

@@

```{r, echo=FALSE,include = FALSE}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Last week summary.

- Discussed PolyMR project with Jonathan. 
  - To start by writing introduction. 
  - Going through paper in-depth.

## Results.

- Univariate MR.

```{r sex_het, echo = FALSE}

tar_load(household_MR_exhaustive_summary)

household_MR_exhaustive_summary <- household_MR_exhaustive_summary %>% mutate(same_traits = case_when(exposure_ID == outcome_ID ~ TRUE, 
                                                                   TRUE ~ FALSE))

DT::datatable(household_MR_exhaustive_summary, width = "100%")

@@

## Questions.

- Expand list of outcomes? 
  - Initially only performed MR with a single trait (i.e. trait_partner ~ trait_index). 
  - Now that we are expanding to two traits in the analysis we could be less stringent with our trait selection.
  - Traits were filtered as follows: 
    1. UKBB phenotypes in SGG database.
    2. Link with Neale summary stats and include only phenotypes which have been processed in both male, female and joint GWAS. 
    3. Evidence of correlation among couples. 
    4. &gt;5 IVs at GW significance. 
    5. No evidence of sex heterogeneity among valid IVs. 
  - Could add more phenotpyes as outcomes

## Next steps.

-

>>>>

---
title: "Update Meeting: 05-07-2021"
author: "Jenny Sjaarda"
date: "2021-07-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

@@

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(household_MR_summary)

@@

# Last week summary.

- Discussed PolyMR project with Jonathan. 
  - To start by writing introduction. 
  - Going through paper in-depth. 
- Next step: run MR between all traits within couples (previously had only looked at single trait at a time).

```{r single_trait_MR, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/single_trait_MR.png", error = FALSE)
```

# Results.

Ran MR between all pairs of traits within couples. Number of tests = $j^2 * 2$, where $j$ is the number of traits. MR models were run in each sex sepeartely (male to female and female to male), which is why we multiply by 2.

```{r two_trait_MR_fig, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
```

The results significant after multiple hypothesis testing can be seen below.

```{r two_trait_MR_results, echo = FALSE}

household_MR_summary.DT <- household_MR_summary %>% dplyr::select(exposure_description, outcome_description, same_trait, exposure_sex, IVW_beta, IVW_pval, IVW_meta_beta, IVW_meta_pval)

num_tests <- dim(household_MR_summary.DT)[1]/2
household_MR_summary.DT.BF <- household_MR_summary.DT %>% filter(IVW_pval < 0.05/num_tests)

DT::datatable(household_MR_summary.DT.BF, width = "100%") %>% formatRound(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval'), digits = 3)

The total number of tests = `r num_tests`, the number significant after adjusting for BF correction = `r dim(household_MR_summary.DT.BF)[1]`.

# Questions.

- Expand list of outcomes? 
  - Initially only performed MR with a single trait (i.e. trait_partner ~ trait_index). 
  - Now that we are expanding to two traits in the analysis we could be less stringent with our trait selection.
  - Traits were filtered as follows: 
    1. UKBB phenotypes in SGG database.
    2. Link with Neale summary stats and include only phenotypes which have been processed in both male, female and joint GWAS. 
    3. Evidence of correlation among couples. 
    4. &gt;5 IVs at GW significance. 
    5. No evidence of sex heterogeneity among valid IVs. 
  - Could add more phenotpyes as outcomes?

# Next steps.

- Calculate MR in same person to finish all analyses in diagram above.

### 2021-09-15 analysis/update_meeting_06_09_2021.Rmd
update all update meetings Rmds

---
title: "Update Meeting: 06-09-2021"
author: "jennysjaarda"
date: "2021-09-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

@@

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

# Model summaries

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
knitr::include_graphics("assets/proxyMR_DAG_MV.png", error = FALSE)

@@

# Questions

1. Impact of meta-analyzing after MR is calculated?

2. Trend across bins: should it be a weighted linear model?

3. Less stringent BF correction: 
    - Calculate correlation matrix of all 131 traits tested. 
    - Perform PCs on the resulting corr matrix. 
    - Number of tests = # of PCs needed until variance explained is > 99.5%.
    - Resulting # of tests: `r num_tests_by_PCs`.

4. Filter by same-person MR or proxyMR?

5. Summed `rho` and `gamma` and compared this to `omega`, for the variance - it's just the sum of these two variances?

6. MV MR: X -> Z -> Y
    - Drop Z's if not significant in MV MR (i.e. `Y ~ X + Z1 + Z2 + Z3 + ...`)
    - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?

# AM summary

```{r AM_summary, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

AM_result <- household_MR_binned_het %>% filter(same_trait)

AM_sig_sex_het <- AM_result %>% filter(sex_het_pval < 0.05/num_tests_by_PCs) %>% dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, sex_het_pval) %>% unique()

AM_sig_age_slope <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_tt_slope <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_age_sex_het <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_tt_sex_het <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

@@

@@

Restricting XY pairs: 
  - If we base it on a significant MR in the same individuals (same MR: $X \rightarrow Y$), the number of tests is: `r dim(standard_MR_summary_BF_sig)[1]`.
  - If we base it on a significant **proxyMR** ($X_{i} \rightarrow Y_{p}$), the number of tests is: `r dim(household_MR_summary_BF_sig)[1]`.

@@

Comparison of omega to rho + gamma instead of just rho or gamma alone. There are more significant results using both.

```{r proxy_MR_update, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

table(proxyMR_comparison_summary$omega_vs_gam_BF_sig_meta)

table(proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta)

table(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta)

@@

t(proxyMR_comparison_summary[205,c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "gam_meta_beta", "rho_meta_beta", "omega_meta_beta", "gam_rho_meta_beta",
                                    "gam_meta_se", "rho_meta_se", "omega_meta_se", "gam_rho_meta_se",
                                   "omega_vs_gam_meta_p", "omega_vs_rho_meta_p", "omega_vs_gam_rho_meta_p")])

>>>>

---
title: "Update Meeting: 06-09-2021"
author: "Jenny Sjaarda"
date: "2021-09-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

@@

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(kableExtra)

@@

# Last meeting summary.

- Need to sum $\rho$ and $\gamma$ together and calculate difference of the ***sum*** from $\omega$ (previously only calculated differences between each pair, not sum).
- Use less stringent BF calculation for AM MR (for calculating trends across bins).

# Model summaries.

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
knitr::include_graphics("assets/MV_trait_model.png", error = FALSE)

@@

# Questions.

1. Impact of meta-analyzing after MR is calculated? 
2. Trend across bins: should it be a weighted linear model? 
3. Less stringent BF correction: 
    - Calculate correlation matrix of all 131 traits tested. 
    - Perform PCs on the resulting corr matrix. 
    - Number of tests = # of PCs needed until variance explained is > 99.5%. 
    - Resulting # of tests: `r num_tests_by_PCs`.  
4. Filter by same-person MR or proxyMR?
5. Summed $\rho$ and $\gamma$ and compared this to `omega`, for the variance - it's just the sum of these two variances? 
6. MV MR: $X \rightarrow  Z \rightarrow Y$
    - Drop Z's if not significant in MV MR (i.e. $Y \sim X + Z_1 + Z_2 + Z_3...$)
    - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?

# AM summary.

A summary of single trait MR within couples (AM MR) can be found [here](AM_MR_summary.html).

@@

Restricting XY pairs:

  - If we base it on a significant MR in the same individuals (same MR: $X \rightarrow Y$), the number of tests is: `r dim(standard_MR_summary_BF_sig)[1]`.
  - If we base it on a significant **proxyMR** ($X_{i} \rightarrow Y_{p}$), the number of tests is: `r dim(household_MR_summary_BF_sig)[1]`.

@@

Comparison of omega to rho + gamma instead of just rho or gamma alone. There are more significant results using both, would have expected there to be less. An example of one such result is shown below.

```{r proxy_MR_update, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

t1 <- as.data.frame(table(proxyMR_comparison_summary$omega_vs_gam_BF_sig_meta))
colnames(t1) <- c("omega_vs_gam_BF_sig_meta", "Frq")

kbl(t1) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

t2 <- as.data.frame(table(proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta))
colnames(t2) <- c("omega_vs_rho_BF_sig_meta", "Frq")

kbl(t2) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

t3 <- as.data.frame(table(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta))
colnames(t3) <- c("omega_vs_gam_rho_BF_sig_meta", "Frq")

kbl(t3) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

@@

sample_dat <- t(proxyMR_comparison_summary[241,c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "gam_meta_beta", "rho_meta_beta", "omega_meta_beta", "gam_rho_meta_beta",
                                    "gam_meta_se", "rho_meta_se", "omega_meta_se", "gam_rho_meta_se",
                                   "omega_vs_gam_meta_p", "omega_vs_rho_meta_p", "omega_vs_gam_rho_meta_p")])
sample_dat <- as.data.frame(sample_dat)
colnames(sample_dat) <- c("Info")

kbl(sample_dat) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

### 2021-09-15 analysis/update_meeting_14_06_2021.Rmd
update all update meetings Rmds

---
title: "Update Meeting: 14-06-2021"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

@@

options(scipen=999)
source("code/settings.R")
load("summary.Rdata")
library(DT)
library(knitr)

@@

```{r, echo=FALSE,include = FALSE}
# You need this code to conduct the magic dependencies attaching...
DT::datatable(matrix())
```

## Last week summary.

### Getting back up to speed.

@@

### Next steps:

@@

### Follow-up questions:

@@

### Responses:

@@

## Results.

### Traits analyzed.

@@

DT::datatable(traits, width = "100%")

@@

### Summary of results

@@

reduced_mr_summary <- calc_mr_sex_het %>% dplyr::select(trait_ID, trait_description, 
                                                        exposure_sex, N_snps,"IVW/Wald_summary","IVW/Wald_pval", 
                                                        "MR_meta_est", "MR_meta_pval", "sex_het")

DT::datatable(reduced_mr_summary, width = "100%")

@@

### Testing discrepancy amongst bins.

@@

mr_summary_age_reduced <- mr_summary_age %>% dplyr::select(trait_ID, exposure_sex, Q_stat, Q_pval)
DT::datatable(mr_summary_age_reduced, width = "100%")

mr_summary_tt_reduced <- mr_summary_tt %>% dplyr::select(trait_ID, exposure_sex, Q_stat, Q_pval)
DT::datatable(mr_summary_tt_reduced, width = "100%")

>>>>

---
title: "Update Meeting: 14-06-2021"
author: "Jenny Sjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

@@

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(DT)
library(knitr)
library(targets)

@@

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(traits_final)
tar_load(household_MR_binned_het)
tar_load(traits_corr2_filled)

# Last week summary.

## Getting back up to speed.

@@

## Next steps:

@@

## Follow-up questions:

@@

## Responses:

@@

# Results.

## Traits analyzed.

@@

traits_final.DT <- traits_final %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_final.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

@@

## Summary of AM MR results

@@

AM_result <- household_MR_binned_het %>% filter(same_trait)
reduced_mr_summary <- AM_result %>% dplyr::select("exposure_ID", "exposure_sex","IVW_beta","IVW_pval", 
                                                        "IVW_meta_beta", "IVW_meta_pval", "sex_het_pval")

reduced_mr_summary.DT <- replace_Neale_ID(reduced_mr_summary, traits_corr2_filled, "exposure_ID", "description")

DT::datatable(reduced_mr_summary.DT, width = "100%") %>% formatRound(columns=c('IVW_beta', 'IVW_pval', 'IVW_meta_beta', 'IVW_meta_pval', 'sex_het_pval'), digits = 3)

@@

## Testing discrepancy amongst bins.

@@

mr_summary_age_reduced <- AM_result %>% dplyr::filter(grouping_var == "age_even_bins") %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, Q_pval, Q_pval_meta)
mr_summary_age_reduced.DT <- replace_Neale_ID(mr_summary_age_reduced, traits_corr2_filled, "exposure_ID", "description")

DT::datatable(mr_summary_age_reduced.DT, width = "100%") %>% formatRound(columns=c('Q_pval', 'Q_pval_meta'), digits = 3)

mr_summary_tt_reduced <- AM_result %>% dplyr::filter(grouping_var == "time_together_even_bins") %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, Q_pval, Q_pval_meta)
mr_summary_tt_reduced.DT <- replace_Neale_ID(mr_summary_tt_reduced, traits_corr2_filled, "exposure_ID", "description")

DT::datatable(mr_summary_tt_reduced.DT, width = "100%") %>% formatRound(columns=c('Q_pval', 'Q_pval_meta'), digits = 3)

### 2021-09-15 analysis/update_meeting_14_07_2021.Rmd
update all update meetings Rmds

---
title: "Update Meeting: 14-07-2021"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

@@

options(scipen=999)
source("code/settings.R")
load("summary.Rdata")
library(DT)
library(knitr)
library(targets)

```{r, echo=FALSE,include = FALSE}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Last week summary.

@@

## Update.

### Reprocessed PHESANT

- For all new phenotype files since starting the project and added them to the pipeline.

```{r new_phenos, echo = FALSE}

@@

### Identified a problem with the Neale summary stats.
- Only affecting the `low_confidence_variant` column. 
- This filter was applied before generating IVs for each summary file that had been downloaded.
- See a description of the issue on [the Neale github issues page](https://github.com/Nealelab/UK_Biobank_GWAS/issues/20).
- Currently in the process of updating the impacted files:
  - Updated Neale GWAS files can be found here: `/data/sgg2/jenny/data/Neale_UKBB_GWAS/neale_files/variant_flag_update`.  
  - Also will recreate the IVs for these files (~70 of our files), under the same name. 
-  As far as I can tell the issue is only in one direction, where some SNPs that were not really low confidence variants (by their definition) were listed as `low_confidence_variant = TRUE`, not the other way around. In other words, the mistake means that we would only be filtering out variants that we should keep.

### Adding PCs to the pipeline.

- Currently running GWAS of first 40 PCs, residualized for `age` and `sex`. Applied inverse normal rank transformation first.  
- Test between partners.

```{r pc_corr, echo = FALSE}

tar_load(PCs_corr)

DT::datatable(PCs_corr, width = "100%")

@@

## Next steps.

@@

$$
\begin{eqnarray*}
	(1) \; X_{i} &=& G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
	(2) \; X_{p} &=& G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
	(3) \; Y_{i} &=& G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
	(4) \; Y_{p} &=& G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
	(5) \; Y_{i} &=& \alpha_j\cdot X_{i} +\epsilon \\
	(6) \; X_{p} &=& \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
	(7) \; Y_{p} &=& \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
	(8) \; Y_{p} &=& \alpha_{k}\cdot X_{i} +\epsilon

\end{eqnarray*}
$$

- Equation (1) and (3) are simply same-person GWAS result (obtained from Neale).
- Equation (2) and (4) are household GWAS results (calculated in-house).
- Equation (5), and ${\alpha_{j}}$, in the causal effect of a trait ${X}$ on ${Y}$ from a standard MR single-person MR.
- Equation (6) and (7), represent assortative mating effects for traits ${X}$ and ${Y}$, respectively.
- Equation (8) ${\alpha_{k}}$ is the result of phenotype ${X}$ in the index, having a causal effect on a phenotype ${Y}$ in the partner. 
- This effect is either due to the product of one of the two:
  - ${\alpha_{j}}$ with ${\alpha_{x, am}}$ 
  - ${\alpha_{j}}$ with ${\alpha_{y, am}}$

>>>>

---
title: "Update Meeting: 14-07-2021"
author: "Jenny Sjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

@@

options(scipen=999)
source("code/settings.R")
library(DT)
library(knitr)
library(targets)
```

# Last week summary.

@@

# Update.

## Reprocessed PHESANT phenotypes.

To include all possible phenotypes, reprocessed PHESANT for all new phenotype files that have been added to SGG databased since starting the project and added them to the pipeline. The list of phenotypes included can be seen below.

```{r new_phenos, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

@@

## Identified a problem with the Neale summary stats.

- Only affecting the `low_confidence_variant` column. 
- This filter was applied before generating IVs for each summary file that had been downloaded.
- See a description of the issue on [the Neale github issues page](https://github.com/Nealelab/UK_Biobank_GWAS/issues/20).
- Currently in the process of updating the impacted files:
  - Updated Neale GWAS files can be found here: `/data/sgg2/jenny/data/Neale_UKBB_GWAS/neale_files/variant_flag_update`.  
  - Also will recreate the IVs for these files (~70 of our files), under the same name. 
-  As far as I can tell the issue is only in one direction, where some SNPs that were not really low confidence variants (by their definition) were listed as `low_confidence_variant = TRUE`, not the other way around. In other words, the mistake means that we would only be filtering out variants that we should keep.

## Adding PCs to the pipeline.

- Currently running GWAS of first 40 PCs, residualized for `age` and `sex`. Applied inverse normal rank transformation first.  
- Test correlation between partners.

```{r pc_corr, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(PCs_corr)
PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr.DT, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)

@@

# Next steps.

@@

$$
\begin{align}
(1) \; X_{i} = G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
(2) \; X_{p} = G_{p}\cdot {\beta_{x, p\rightarrow p} } + \epsilon \\
(3) \; X_{p} = G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
(4) \; Y_{i} = G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
(5) \; Y_{p} = G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
(6) \; Y_{i} = \alpha_{x_{i} \rightarrow y_{i}}\cdot X_{i} +\epsilon \\
(7) \; Y_{p} = \alpha_{x_{p} \rightarrow y_{p}}\cdot X_{p} +\epsilon \\
(8) \; X_{p} = \alpha_{x_{i} \rightarrow x_{p}}\cdot X_{i{}} +\epsilon \\
(9) \; Y_{p} = \alpha_{y_{i} \rightarrow y_{p}}\cdot Y_{i{}} +\epsilon \\
(10) \; Y_{p} = \alpha_{x_{i} \rightarrow y_{p}}\cdot X_{i} +\epsilon
\end{align}
$$

- Equation (1), (2), and (4) are simply same-person GWAS result (obtained from Neale).
- Equation (3) and (5) are household GWAS results (calculated in-house).
- Equation (6) and (7) (${\alpha_{x_{i} \rightarrow y_{i}}}$ and $\alpha_{x_{p} \rightarrow y_{p}}$), in the causal effect of a trait ${X}$ on ${Y}$ from a standard MR single-person MR, sex specific.
- Equation (8) and (9), represent assortative mating effects for traits ${X}$ and ${Y}$, respectively ($\alpha_{x_{i} \rightarrow x_{p}}$ and $\alpha_{y_{i} \rightarrow y_{p}}$).
- Equation (10) $\alpha_{x_{i} \rightarrow y_{p}}$ is the result of phenotype ${X}$ in the index, having a causal effect on a phenotype ${Y}$ in the partner. 
    - This effect is either due to:
        (1) The product of $\alpha_{x_{i} \rightarrow x_{p}}$ with $\alpha_{x_{p} \rightarrow y_{p}}$, or 
        (2) The product of $\alpha_{x_{i} \rightarrow y_{i}}$ with $\alpha_{y_{i} \rightarrow y_{p}}$, or
        (3) An additional 3rd (or 4th, etc.) variable which amplifies the effect from ${X_{i}}$ to ${Y_{p}}$ that doesn't simply act through a combination of assortative mating and a causal effect of $X$ on $Y$.
